Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-26 chatbot_classification/aaai_2021/paper.tex
modified:   chatbot_classification/aaai_2021/paper.pdf   modified:   chatbot_classification/aaai_2021/paper.tex

\begin{figure}[htbp]
\centering 
\subfigure[classification model]{
\label{Fig.1.sub.1}
\includegraphics[scale=0.7]{picture/picture1.pdf}}
\subfigure[similarity model]{
\label{Fig.1.sub.2}
\includegraphics[scale=0.55]{picture/picture2.pdf}}
\caption{Two popular structures of building task-specific conversational chatbot}
\label{Fig.1.main}
\end{figure}

One of the most popular approach among existing work was text classification model based on neural network. Since the length of the text is quite short, many previous work\cite{wen2016network} choose neural network such as convolutional neural networks (CNNs)\cite{kim2014convolutional,zhang2015character,conneau2016very}
or long short term memory networks
(LSTMs)\cite{mousa2017contextual,liu2016recurrent} to accomplish the task of extracting semantic feature from limited amount of words. A common model structure is adding a softmax classifier to the top of the neural network. Another common approach is text classification neural network based on label-word joint embedding, such as LEAM\cite{wang2018joint}. This approach involves label embedding when constructing the text-sequence representation, which poses requirements on label information. In our cases, since we have large amount of class labels without being well-defined, this approach can do only a few help to the low-resource chatbot scenario. Afterwards, pre-trained language models on large corpus like BERT\cite{devlin2018bert} and
RoBERTa\cite{liu2019roberta} has  been proven more powerful in solving many NLP
tasks      including      short-text      classification\cite{madabushi2020cost}.   Especially   for  few-shot  scenarios\cite{yu2018diverse}, pre-trained  model  based on transformers\cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends  to  do more help to reducing the negative effects brought by scarcity of training data.

Another popular approach among previous work was based on sentence-pair model. The motivation of this approach was started from the idea of Information Retrival (IR) based chatbot\cite{jafarpour2010filter, leuski2011npceditor}. Having a Q-A (Question-Answer) pairs dataset and user query Q, the IR based conversational system will look up in the Q-A dataset for the pair (Q', A') that best matches query Q through semantic analysis and returns A' as the answer to Q\cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large corpus of semantic similarity identification task can be applied for being used as a tool to identify the class with highest semantic similarity to customer's query. A common model structure of pre-trained\cite{devlin2018bert} sentence-pair model is based on multiple cross-attention mechanism\cite{barkan2020scalable}, shown in Fig. \ref{Fig.1.sub.2}. Due to the fact that many experiment results have shown that RoBERTa\cite{liu2019roberta} is an improved  version  of  BERT, and has achieved amazing results on both text classification and sentence-pairs semantic similarity tasks, we choose RoBERTa as an important baseline approach in this paper.

Despite the success of these 2 approaches, they still have some limitations in task-specific chatbot scenario. As for the text classification model approach, it's quite hard to use data augmentation method to solve the data insufficiency challenge since the it's usually unfeasible to get large amount of domain-specific data when facing a new task. With respect to the sentence-pair model approach, though we can obtain large amount of data in semantically duplicate sentence pair identification domain for transfer learning\cite{sun2019fine}, it's still quite hard to perform well on task-specific dataset because the training objective of sentence-pair model is semantic similarity, which is slightly different from the target of classification task. That is to say, there always exists some intent classes which cannot be distinguished by each other merely depending on semantic similarity. For example, we have 2 user query saying, A: What should I do if I want to change my password? B: What is the modification rule if I want to change my password? In this case, sentence A is semantically similar to B, since they both express the 
desire for changing password. However, it's still reasonable to classify them into 2 intent classes, since A is asking for the method for changing password, while B is asking for the rule to follow when creating a new password.

\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.48]{picture/picture3}
\par\end{centering}
\caption{\textbf{Network Structure of 2-stage SFC:} a joint system of multi-task for both text classification task and sentence-pair semantic similarity task.}
\label{fig:ver.1.3}
\end{figure}

@@

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
\multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & - & -\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\caption{We show the performance of SFC from different settings of hyperparameters for $K$, the candidate class number, and $P$, the sample number for each class.}

>>>>

  \begin{figure}[htbp]
    \centering 
    \subfigure[classification model]{
      \label{Fig.1.sub.1}
      \includegraphics[scale=0.7]{picture/picture1.pdf}
    }
    \subfigure[similarity model]{
      \label{Fig.1.sub.2}
      \includegraphics[scale=0.55]{picture/picture2.pdf}
    }
    \caption{Two popular structures of building task-specific conversational chatbot}
    \label{Fig.1.main}
  \end{figure}

  One  of  the most popular approach among existing work was text classification
  model  based  on  neural network. Since the length of the text is quite short,
  many   previous   work\cite{wen2016network}  choose  neural  network  such  as
  convolutional                          neural                         networks
  (CNNs)\cite{kim2014convolutional,zhang2015character,conneau2016very}  or  long
  short  term memory networks (LSTMs)\cite{mousa2017contextual,liu2016recurrent}
  to  accomplish  the task of extracting semantic feature from limited amount of
  words.  A  common model structure is adding a softmax classifier to the top of
  the  neural  network.  Afterwards, pre-trained language models on large corpus
  like   BERT\cite{devlin2018bert}  and  RoBERTa\cite{liu2019roberta}  has  been
  proven   more   powerful  in  solving  many  NLP  tasks  including  short-text
  classification\cite{madabushi2020cost}.      Especially      for      few-shot
  scenarios\cite{yu2018diverse},       pre-trained      model      based      on
  transformers\cite{vaswani2017attention},   shown  in  Fig.  \ref{Fig.1.sub.1},
  tends  to do more help to reducing the negative effects brought by scarcity of
  training data.

  Another  common  approach  is  text  classification  neural  network  based on
  label-word     joint     embedding,    such    as    LEAM\cite{wang2018joint},
  MTLE\cite{zhang2017multi}   and   EXAM\cite{du2019explicit}.   This   approach
  involves  label  embedding when constructing the text-sequence representation.
  Take  LEAM  as  an  example,  the  text classification model is implemented by
  jointly    embedding    the    word    and    label   in   the   same   latent
  space\cite{wang2018joint}. Afterwards, the text representation are constructed
  based  on  text-label  compatibility through attention mechanism. In this way,
  additional  sources  of  semantic  information  from class labels can be fully
  leveraged, which is quite helpful especially under few-shot scenario.

  The  third  popular  approach  among  previous work was based on sentence-pair
  model.  The  motivation  of  this  approach  was  started  from  the  idea  of
  Information     Retrival    (IR)    based    chatbot\cite{jafarpour2010filter,
  leuski2011npceditor}.  Having  a  Q-A (Question-Answer) pairs dataset and user
  query  Q,  the  IR based conversational system will look up in the Q-A dataset
  for  the pair (Q', A') that best matches query Q through semantic analysis and
  returns   A'   as   the  answer  to  Q\cite{mnasri2019recent}.  In  this  way,
  sentence-pair  model  pre-trained  on  large  corpus  of  semantic  similarity
  identification  task  can  be applied for being used as a tool to identify the
  class  with  highest  semantic  similarity to customer's query. A common model
  structure  of pre-trained\cite{devlin2018bert} sentence-pair model is based on
  multiple  cross-attention  mechanism\cite{barkan2020scalable},  shown  in Fig.
  \ref{Fig.1.sub.2}.  Due  to  the  fact that many experiment results have shown
  that  RoBERTa\cite{liu2019roberta}  is  an  improved  version of BERT, and has
  achieved  amazing  results  on  both  text  classification  and sentence-pairs
  semantic similarity tasks, we choose RoBERTa as an important baseline approach
  in this paper.

  Despite the success of these 3 approaches, they still have some limitations in
  task-specific chatbot scenario. As for the text classification model approach,
  it's   quite   hard  to  use  data  augmentation  method  to  solve  the  data
  insufficiency  challenge since the it's usually unfeasible to get large amount
  of  domain-specific  data  when  facing  a  new  task. As for label-word joint
  embedding  approach,  it  poses  noticeable requirements on label information.
  However,  in  our  cases,  since  we have large amount of class labels without
  being well-defined for question answering chatbot, the only thing we can do is
  to  choose  one  of the user utterance as the class label to represent the key
  information  of  the  whole class. In this way, the label will become not only
  prolix  but  also  biased,  which  will  bring  some  negative effect to label
  embedding.  With  respect  to  the sentence-pair model approach, though we can
  obtain   large   amount  of  data  in  semantically  duplicate  sentence  pair
  identification  domain  for  transfer  learning\cite{sun2019fine},  it's still
  quite  hard  to  perform  well  on  task-specific dataset because the training
  objective  of  sentence-pair  model  is semantic similarity, which is slightly
  different from the target of classification task. That is to say, there always
  exists  some intent classes which cannot be distinguished by each other merely
  depending on semantic similarity. For example, we have 2 user query saying, A:
  What  should I do if I want to change my password? B: What is the modification
  rule if I want to change my password? In this case, sentence A is semantically
  similar  to  B,  since  they  both  express  the desire for changing password.
  However, it's still reasonable to classify them into 2 intent classes, since A
  is asking for the method for changing password, while B is asking for the rule
  to follow when creating a new password.

  \begin{figure}[htbp]
    \begin{centering}
      \includegraphics[scale=0.48]{picture/picture3}
      \par
    \end{centering}
    \caption{
      \textbf{Network  Structure  of  2-stage SFC:} a joint system of multi-task
      for  both  text  classification task and sentence-pair semantic similarity
      task.
    }
    \label{fig:ver.1.3}
  \end{figure}

@@

      \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline 
        \multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
        \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
        \hline
      \end{tabular}
      \par
    \end{centering}
    \caption{
      We  show the performance of SFC from different settings of hyperparameters
      for  $K$,  the candidate class number, and $P$, the sample number for each
      class.
    }
